Short-term synaptic and neuronal plasticity act in synergy for deviance detection in spiking networks

### 2022-12-15 paper-1/manuscript.tex
Editing

% adding STD: What happens?
As noted previously, short-term depression alone exerted no deviance detection effect, but its addition did enhance the deviance detection effect established by TA. To understand the basis of this apparent synergy, we contrasted the TA only condition examined in the previous section against the same networks with both TA and STD. On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (data not shown). However, this reduction was not uniform: Target responses were reduced most in the std sequence, and least in the dev sequence, this difference leading directly to the greater DD and SSA indices reported in Figure \ref{fig3}.

% Why does dXA increase?:
To understand the reason for this difference, we first focused on the target response and asked how and why the target response contrast (dev - msc) changed as STD was added to the model. Firstly, as shown in Figure \ref{fig7}a, the target response was greater in dev than msc in both the full and the ablated model (indicating a positive DDI). In most neurons, the contrast was greater in the full model, such that the difference (full - ablated) was positive, as indicated by the largely orange plot in the right column.
Following the logic established in the previous section, we may expect a greater target response in dev to be caused by lower $V_{TA}$ relative to msc. Indeed, comparing the contrast (dev - msc) in $V_{TA}$ before target trials across models (Figure \ref{fig7}b), we see that, while the contrast is negative (i.e., lower in dev) in both models, the magnitude of the contrast is greater in the full model. For clarity, the difference is therefore negative, as indicated by the largely blue plot in the right column.
Lastly, we trace this one step further back in Figure \ref{fig7}c: Again as established in the previous section, a greater contrast in $V_{TA}$ may be explained by a similarly enhanced contrast (dev - msc) in the non-target response, which is exactly what we found: The non-target response was smaller in dev than msc by a greater margin in the full model, as indicated by the largely blue plot in the right column.

This pattern is perhaps simpler to grasp in the aggregate across networks and stimuli. Firstly, the contrast (dev - msc) in the target response (Figure \ref{fig7}d, left) was positive in both models, but the full model contrast was greater than the ablated model contrast (t = 2.99, p = 0.00168; one-sided t-test, n = 120). Referring to the previous section, we locate the cause for this in an enhanced contrast in $V_{TA}$, as shown in Figure \ref{fig7}e; the contrast (dev - msc) in the full model was more negative than that in the ablated model (t = -3.49, p = 0.000342). Lastly, returning to Figure \ref{fig7}d (right), we see that the driver of $V_{TA}$, namely non-target response size, also showed an enhanced contrast (dev - msc) in the full model relative to the ablation (t = -3.19, p = 0.000921). In other words, the reason for the larger target dev response, relative to msc, is the smaller non-target dev response, relative to msc.

>>>>

As noted previously, short-term depression alone exerted no deviance detection effect, but its addition did enhance the deviance detection effect established by TA. To understand the basis of this apparent synergy, we contrasted the TA only condition examined in the previous section against the same networks with both TA and STD. On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (data not shown). However, this reduction was not uniform: Target responses were reduced most in the std sequence, and least in the dev sequence, this difference leading directly to the greater DD and SSA indices reported in Figure \ref{fig3}.

To understand the reason for this difference, we first focused on the target response and asked how and why the target response contrast (dev - msc) changed as STD was added to the model. Firstly, as shown in Figure \ref{fig7}a, the target response was greater in dev than msc in both the full and the ablated model (indicating a positive DDI). In most neurons, the contrast was greater in the full model, such that the difference (full - ablated) was positive, as indicated by the largely orange plot in the right column.
Following the logic established in the previous section, we may expect a greater target response in dev to be caused by lower $V_{TA}$ relative to msc. Indeed, comparing the contrast (dev - msc) in $V_{TA}$ before target trials across models (Figure \ref{fig7}b), we see that, while the contrast is negative (i.e., lower in dev) in both models, the magnitude of the contrast is greater in the full model. For clarity, the difference is therefore negative, as indicated by the largely blue plot in the right column.
Lastly, we trace this one step further back in Figure \ref{fig7}c: Again as established in the previous section, a greater contrast in $V_{TA}$ may be explained by a similarly enhanced contrast (dev - msc) in the non-target response, which is exactly what we found: The non-target response was smaller in dev than msc by a greater margin in the full model, as indicated by the largely blue plot in the right column.

This pattern is perhaps simpler to grasp in the aggregate across networks and stimuli. Firstly, the contrast (dev - msc) in the target response (Figure \ref{fig7}d, left) was positive in both models, but the full model contrast was greater than the ablated model contrast (t = 2.99, p = 0.00168; one-sided t-test, n = 120). Referring to the previous section, we locate the cause for this in an enhanced contrast in $V_{TA}$, as shown in Figure \ref{fig7}e; the contrast (dev - msc) in the full model was more negative than that in the ablated model (t = -3.49, p = 0.000342). Lastly, returning to Figure \ref{fig7}d (right), we see that the driver of $V_{TA}$, namely non-target response size, also showed an enhanced contrast (dev - msc) in the full model relative to the ablation (t = -3.19, p = 0.000921). In other words, the reason for the larger target dev response, relative to msc, is the smaller non-target dev response, relative to msc; furthermore, since the relationships established in the previous section remained largely unchanged, this effect appeared to be mediated primarily through TA.
